Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating subgenome assignments in maize

### 2017-08-11 R/SubGenome Analysis.Rmd
Provided support for a basic GO analysis
Cleaner data slices, parsed Citation field, but needs more cleanup, we have some unused and obsolete objects generated here.

Can do a basic GO analysis on the GO terms themselves, but not yet on their assignment to genes.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(ggplot2)
library(tidyr)
library(dplyr)
```

@@

# SubGenomes
***
In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size.

```{r reportSubGenomes, echo=FALSE, results="asis"}
# library(venneuler)
# i <- intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
# t <- setdiff(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
# p <- setdiff(subgenome.truth, subgenome.chromosomes.test) %>% nrow()
# vd <- venneuler(c(Paper=p, Test=t, "Paper&Test"=i))
# vd$labels <- c(
#   paste("Paper\n",p+i),
#   paste("Test\n",t+i),
#   paste("Both\n",i)
# )
# plot(vd)
```

## Compared to the original paper.
How many chromosomes with synteny to sorghum were correctly placed in subgenome 1 or subgenome 2? Items in "Paper" only are missed subgenome assignments.  Items in "Test" only are false assignments.  Items in center are correct. (this is an unfair comparison if I sorted by hand)

```{r compareResultsToPaper, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
library(VennDiagram)
# i <- intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
# t <- setdiff(subgenome.chromosomes.test, subgenome.truth) %>% nrow()
# p <- setdiff(subgenome.truth, subgenome.chromosomes.test) %>% nrow()
subgenome.chromosomes.test <- 
  subgenome.chromosomes %>% 
  mutate(chr1=as.numeric(regmatches(org_chr1, regexpr("\\d*$",org_chr1)))) %>%
  mutate(chr2=as.numeric(regmatches(org_chr2, regexpr("\\d*$",org_chr2)))) %>%
  select(chr1, chr2, subgenome) %>%
  subset(subgenome == "sub1" | subgenome == "sub2")

draw.pairwise.venn(
  area1 = subgenome.truth%>%nrow(), 
  area2 = subgenome.chromosomes.test%>%nrow(), 
  cross.area = intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow(), 
  category = c("Paper", "Test"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)
```

# GOTerms
***

***

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(ggplot2)
library(tidyr)
library(dplyr)
library(VennDiagram)
```

@@

# SubGenomes
***
In order to separate subgenome 1 from subgenome 2, we implement a greedy algorithm to collect non-overlapping (when projected on sorghum) syntenic blocks by size. (TODO)

```{r reportSubGenomes, echo=FALSE, results="asis"}
#Positions on each chr with sub 1 or sub 2
#Total genes for each subgenome by chr and overall
```

## Compared to Paper
How many chromosomes with synteny to sorghum were correctly placed in subgenome 1 or subgenome 2? Items in "Paper" only are missed subgenome assignments.  Items in "Test" only are false assignments.  Items in center are correct. (this is an unfair comparison if I sorted by hand)

```{r compareResultsToPaper_chromosomes, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
subgenome.chromosomes.test <- 
  subgenome.chromosomes %>% 
  mutate(chr1=as.numeric(regmatches(org_chr1, regexpr("\\d*$",org_chr1)))) %>%
  mutate(chr2=as.numeric(regmatches(org_chr2, regexpr("\\d*$",org_chr2)))) %>%
  select(chr1, chr2, subgenome) %>%
  subset(subgenome == "sub1" | subgenome == "sub2")

grid.newpage()
draw.pairwise.venn(
  area1 = subgenome.truth %>% nrow(), 
  area2 = subgenome.chromosomes.test %>% nrow(), 
  cross.area = intersect(subgenome.chromosomes.test, subgenome.truth) %>% nrow(), 
  category = c("Paper", "Test"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)
```

Genes placed in the "correct" subgenome (TODO)

```{r compareResultsToPaper_genes, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=3, fig.align='center'}
#need to import "truth" and compare
#visualize as venn
```

# GOTerms
***
GO Term types by EVCode

```{r GOTerms_by_EVCode, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
Type <- c("EV-EXP","EV-COMP")
Count <- c(goAnnotations.mutated %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
      goAnnotations.mutated %>% filter(grepl("EV-COMP", EVCode)) %>% nrow())
df = data.frame(Type,Count)
ggplot(data=df, aes(x=Type, y=Count)) +
  geom_bar(stat="identity", fill="steelblue") +
  geom_text(aes(label=Count), vjust=-0.3, size=3.5)
```

GO Term by subgenome (All EV-Codes)

```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
# All EV-Codes, Unique GO Terms
grid.newpage()
draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% select("GO Term") %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% select("GO Term") %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% select("GO Term"), goAnnotations.sub2 %>% select("GO Term")) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)

#EXP only, Unique GO Terms
grid.newpage()
draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term") %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term"), 
                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select("GO Term")
                         ) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)
```

GOTerm assignments to genes by subgenome (All EV-Codes)
(broken - doesn't take into account homeologous genes - try replacing equivalent genes from homeolog list)

```{r GOTerms_assignments_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
# All EV-Codes, all assignments
grid.newpage()
draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% nrow(),
  area2 = goAnnotations.sub2 %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% select(gene2, "GO Term"), goAnnotations.sub2 %>% select(gene2, "GO Term")) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)

#EXP only, all assignments
grid.newpage()
draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
  area2 = goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term"), 
                         goAnnotations.sub2 %>% filter(grepl("EV-EXP", EVCode)) %>% select(gene2, "GO Term")
                         ) %>% nrow(), 
  category = c("SubGenome 1", "SubGenome 2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0
)
```

# Wrapup
***
